Cheminformatics: The Computer Science of Chemical Discovery

### 2011-11-09 paper.tex
Remove statement that 3d info is completely lost. Replace by other statements on the topic.

Graph mining algorithms can be applied to this type of annotated graph
(employing large databases of thousands of molecules), to discover
patterns in large-scale datasets.  Many successful applications of
this concept have been published \cite{wegner2006,horst2009}. Still,
this ``annotated graph'' representation is far from being the end of
the story. Molecules are three-dimensional entities, as represented in
the bottom left hand corner of the figure, and the precise 3D shape is
relevant in many situations (although not in others).  Nevertheless, the
3D information is lost when one only considers the graph of a
molecule. To extend the concept of three-dimensionality further, in
many biological systems the \emph{surface properties} of the molecule
are what appear to be relevant, as shown in the bottom right hand
corner.  Hence, while graph mining on chemicals is certainly a
promising avenue to explore, one needs to keep in mind that not all
molecular properties are encoded in the graph (for example, neither
rotational states, nor different charge states are encoded) and thus
often other ``descriptors'' need to be used to capture the essentials of
a chemical system \cite{bender2004}. Hence, one of the difficulties of a cheminformatician is that the equivalence of
particular atoms (or sets of atoms) is usually not known from the onset but it rather depends on the particular system
considered---involving a large amount of trial-and-error in the choice of a ``proper'' description for molecules in a
particular case.

>>>>

Graph mining algorithms can be applied to this type of annotated graph
(employing large databases of thousands of molecules), to discover
patterns in large-scale datasets.  Many successful applications of
this concept have been published \cite{wegner2006,horst2009}. Still,
this ``annotated graph'' representation is far from being the end of
the story. Molecules are three-dimensional entities, as represented in
the bottom left hand corner of the figure, and the precise 3D shape is
relevant in many situations (although not in others). Although the
graph implicitly encodes the 3D structure of a molecule (when combined
with knowledge about favored bond angles and distances), many
different low energy 3D structures, or conformers, may be consistent
with the same graph structure. In addition, the 3D structure may have
a geometrical arrangement of features (such as the presence of a
right-handed helix) which cannnot be encoded in the graph.

To extend the concept of three-dimensionality further, in
many biological systems the \emph{surface properties} of the molecule
are what appear to be relevant, as shown in the bottom right hand
corner.  Hence, while graph mining on chemicals is certainly a
promising avenue to explore, one needs to keep in mind that not all
molecular properties are encoded in the graph (for example, neither
rotational states, nor different charge states are encoded) and thus
often other ``descriptors'' need to be used to capture the essentials of
a chemical system \cite{bender2004}. Hence, one of the difficulties of a cheminformatician is that the equivalence of
particular atoms (or sets of atoms) is usually not known from the onset but it rather depends on the particular system
considered---involving a large amount of trial-and-error in the choice of a ``proper'' description for molecules in a
particular case.
